Mathematical Models of Delta-Notch Signalling

### 2025-03-15 paper/paper.tex
Add abstract to paper.tex

>>>>

The Delta-Notch signalling pathway plays a critical role in determining the cell fate of initially homogeneous cells in diverse biological systems. Early deterministic models of Delta-Notch interactions, such as that by Collier et al. (1996), achieved cell differentiation in linear and 2D cell arrays. In this paper, we develop a simple – yet more biologically realistic – ODE model of Delta-Notch signalling and compare the deterministic behaviour with the Collier et al. model. We derive a probabilistic framework via Kolmogorov forward equations, demonstrating that, in expectation, it converges to our ODE model under appropriate limiting conditions. By comparing deterministic ODEs, stochastic ODEs, and Gillespie simulations, we show that natural perturbations in receptor-ligand binding dynamics are sufficient to produce differentiation, even from identical initial conditions. Numerical experiments across parameter regimes highlight the robustness of this behaviour and bifurcation analysis further reveals the parameter-dependent bistability.
